---
title: "02_filtering"
author: "Jack Hardwick, Somdutta Dhir, Angela Simeone"
date: "2024-03-25"
output: html_document
---

This script performs filtering on the individual replicates. 
The output (which can be parsed to downstream scripts) is used only to compare replicates and for the GEO deposition, as the main analysis is performed on merged replicates. 

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
# define spike-ins
spike_ins <- c("H1", "H2", "Lambda", "pUC19", "SQ4mC", "SQ2hmC", "SQC")
# define hmC states
hmC_states <- c("HC", "CH", "HM", "MH", "HH")

# Read config file where paths are defined
config <- read.table("config.txt", 
                     sep="=", 
                     comment.char="#", 
                     col.names=c("key", "value"))

# Convert contents to individual variables
config_list <- as.list(setNames(config$value, config$key)) %>%
  list2env(envir = .GlobalEnv)
```

# SCoTCH-seq: filtering and binning, and generation of bed, bedgraph and bigwig files

The following script takes the output from the processing pipeline of both replicates (`./data/pipeline_output/CpG_state_quants_*`) and does the following:
* Removes spike-ins and unwanted chromosomes (e.g. ChrM, chrUn_*)
* Filters for minimum depth (≥ 5x per CpG site)
* Error filtering
* Blacklist filtering

### Import data, remove spike-ins and unwanted chromosomes, remove CpG sites with depth < 5
```{r}
# Import CpG-state info (raw pipeline output)
# and filter out unwanted chromosomes

# Import data and remove unwanted chromosomes
quants_e14_filt <-
  read_tsv(paste0(path_quants_rep1, rep1)) %>%
  subset(!grepl("chrUn_|random|chrM", Chr))

# Rename column
colnames(quants_e14_filt) <-
  c("Chr", "CpG_ID", "CpG_State", "Count", "pct_prefiltering")

# Assign strands to CpG states (needed for some programmes).
# CC, MM, HH are arbitrarily denoted '+'
quants_e14_filt <- quants_e14_filt %>%
  mutate(start = CpG_ID) %>%
  mutate(end = CpG_ID + 1) %>%
  mutate(strand = case_when(CpG_State == "MC" ~ "+",
                            CpG_State == "CM" ~ "-",
                            CpG_State == "HC" ~ "+",
                            CpG_State == "CH" ~ "-",
                            CpG_State == "HM" ~ "+",
                            CpG_State == "MH" ~ "-",
                            TRUE ~ "+"))

# Create unique ID for each CpG site by merging chromosome with CpG start coord
quants_e14_filt$CpG_ID <-
  paste0(quants_e14_filt$Chr,
         sep = "_",
         quants_e14_filt$start)

# Error filtering:
## Get IDs of mESC CpG sites containing multiple CpG-state calling errors
CpGs_with_errors <-
  quants_e14_filt %>%
  filter(!Chr %in% spike_ins & CpG_State == "Error" & Count > 1) %>%
  select(CpG_ID)

## Remove any mESC CpG sites with multiple CpG-state errors from the data
quants_e14_filt <-
  anti_join(quants_e14_filt,
            CpGs_with_errors,
            join_by(CpG_ID))

## Remove errors from filtered data
quants_e14_filt <-
  quants_e14_filt %>%
  filter(CpG_State != "Error")

# Determine depth at each CpG site after error filtering
CpG_depth <- quants_e14_filt %>%
  group_by(Chr, CpG_ID) %>%
  reframe(depth = sum(Count))

# Merge depth with CpG_state_quants
quants_e14_filt <-
  left_join(quants_e14_filt, CpG_depth, by = c("Chr", "CpG_ID"))

# Calculate percentage
quants_e14_filt <-
  quants_e14_filt %>%
  mutate(pct = 100 * Count / depth)

# Set minimum depth to 5x and reorder
quants_e14_filt_depth_over_5x <-
  quants_e14_filt %>%
  filter(depth >= 5) %>%
  dplyr::select(Chr,
                start,
                end,
                CpG_ID,
                CpG_State,
                pct,
                Count,
                depth,
                strand) %>%
  write_tsv(paste0(intermed_files, "quants_e14_filt_5x_rep1.bed"))
```

### Blacklist filtering (cite blacklist source https://doi.org/10.1038/s41598-019-45839-z)
```{zsh engine.opts='-i'}
conda activate doubleshot

# Load config file
source config.txt

# Change directory
cd $intermed_files

# Create new file containing column names
head -n 1 quants_e14_filt_5x_rep1.bed > quants_e14_filt_5x_rep1_colnames.txt

# Remove CpG sites intersecting with mm10 blacklist and add column names to output
tail -n +2 quants_e14_filt_5x_rep1.bed | bedtools intersect -v -a stdin -b ${misc_files}mm10-blacklist.v2.bed | bedtools sort -i stdin | cat quants_e14_filt_5x_rep1_colnames.txt - > quants_e14_5x_blacklist_filt_rep1.bed

# Remove unwanted input files
#rm quants_e14_filt_5x.bed 
#rm quants_e14_filt_5x_colnames.txt
```

### Extract coordinates of CpG sites covered by merged replicates (used for enhancer analysis)
```{r}
read_tsv(paste0(intermed_files,
         "quants_e14_5x_blacklist_filt_rep1.bed")) %>%
  dplyr::select(Chr, start, end) %>%
  distinct() %>%
  write_tsv(paste0(intermed_files,
                   "quants_e14_5x_blacklist_filt_CpG_coords_rep1.bed"),
            col_names = FALSE)
```

