---
title: "call_rate_matrix"
output: html_document
date: "2023-11-22"
---
# The following script generates the call-rate (confusion) matrices for individual and merged replicates. 


## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reshape2)
library(viridis)
options(scipen = 999)

# Read config file where paths are defined
config <- read.table("config.txt", 
                     sep="=", 
                     comment.char="#", 
                     col.names=c("key", "value"))

# Convert contents to individual variables
config_list <- as.list(setNames(config$value, config$key)) %>%
  list2env(envir = .GlobalEnv)
```

## Create output directory
```{bash}
source config.txt
mkdir -p $call_rate_output
```

## Synthetic spike-ins
```{r}
# Define synthetic spike-ins
synth_spike <- c("H1", "H2", "SQ4mC", "SQ2hmC", "SQC")

# Import CpG states from synthetic spike-ins
all_reps_synth_spike <- 
  read_tsv(paste0(intermed_files, 
                  "quants_spikeins_reps_merged.tsv")) %>%
  filter(Chr %in% synth_spike)

# Separate dataframes by replicate...

# Rep 1 (JSH9)
JSH9_synth_spike <- all_reps_synth_spike %>%
  select(Chr, CpG_ID, CpG_State, n.JSH9)
colnames(JSH9_synth_spike)[4] <- "Count"

# Rep 2 (JSH11)
JSH11_synth_spike <- all_reps_synth_spike %>%
  select(Chr, CpG_ID, CpG_State, n.JSH11)
colnames(JSH11_synth_spike)[4] <- "Count"

# Merged reps
merged_reps_synth_spike <- all_reps_synth_spike %>%
  select(Chr, CpG_ID, CpG_State, n_merged)
colnames(merged_reps_synth_spike)[4] <- "Count"

# Create list of dataframes
list_all_reps <- list(JSH9_synth_spike, 
                      JSH11_synth_spike, 
                      merged_reps_synth_spike)

# Define function for initial processing
fn_initial_processing <- function(df) {
  df %>%
# 1. Remove errors
    filter(CpG_State != "Error") %>%
    group_by(Chr, CpG_ID) %>%
# 2. Calculate depth:
    mutate(depth = sum(Count)) %>%
    ungroup() %>%
# 3. Filter to remove small fraction of probable alignment errors
    filter(depth > 20) %>%
# 4. Caclulate percentage of each CpG state called
# 5. Correct CpG IDs ('NNNN' was added to spike-in references
#    prior to alignment to resolve issue aligning to short references)
    mutate(pct_adjust = 100 * Count / depth, CpG_ID = CpG_ID - 4)
}

# Apply above function to list of df's
initial_processed_list <- map(list_all_reps, fn_initial_processing)

# Fn to sum percentages (which should add up to 100...)
fn_sanity_check <- function(df) {
  df %>%
    group_by(Chr, CpG_ID) %>%
    reframe(sum(pct_adjust))
}

# Apply sanity check to list
sanity_checked_list <- map(initial_processed_list, fn_sanity_check)
# Passed check

# Function to add ground truth info to df
fn_add_ground_truths <- function(df) {
  df %>%
    mutate(ground_truth = case_when((Chr == "H1" & CpG_ID == 13) ~ "CC",
                                  (Chr == "H1" & CpG_ID == 28) ~ "HH",
                                  (Chr == "H1" & CpG_ID == 44) ~ "CC",
                                  (Chr == "H1" & CpG_ID == 58) ~ "MM",
                                  (Chr == "H2" & CpG_ID == 19) ~ "CC",
                                  (Chr == "H2" & CpG_ID == 23) ~ "HM",
                                  (Chr == "H2" & CpG_ID == 28) ~ "CC",
                                  (Chr == "H2" & CpG_ID == 51) ~ "CH",
                                  (Chr == "H2" & CpG_ID == 60) ~ "CC",
                                  (Chr == "H2" & CpG_ID == 74) ~ "MC",
                                  (Chr == "SQ2hmC" & CpG_ID == 22) ~ "HH",
                                  (Chr == "SQ2hmC" & CpG_ID == 38) ~ "CC",
                                  (Chr == "SQ2hmC" & CpG_ID == 59) ~ "HC",
                                  (Chr == "SQ4mC" & CpG_ID == 21) ~ "CM",
                                  (Chr == "SQ4mC" & CpG_ID == 38) ~ "CC",
                                  (Chr == "SQ4mC" & CpG_ID == 59) ~ "MC",
                                  (Chr == "SQ4mC" & CpG_ID == 67) ~ "MM",
                                  (Chr == "SQC" & CpG_ID == 13) ~ "CC",
                                  (Chr == "SQC" & CpG_ID == 19) ~ "CC",
                                  (Chr == "SQC" & CpG_ID == 22) ~ "CC",
                                  (Chr == "SQC" & CpG_ID == 28) ~ "CC",
                                  (Chr == "SQC" & CpG_ID == 32) ~ "CC",
                                  (Chr == "SQC" & CpG_ID == 45) ~ "CC",
                                  (Chr == "SQC" & CpG_ID == 52) ~ "CC",
                                  (Chr == "SQC" & CpG_ID == 59) ~ "CC", ))
}

gt_list <- map(initial_processed_list, fn_add_ground_truths)

# Generate mean of ground truth and CpG-state calls and fill in missing values
fn_mean_calls <- function(df) {
  df %>%
    group_by(ground_truth, CpG_State) %>%
# 1. Generate mean calls
    reframe(call_rate = mean(pct_adjust)) %>%
# 2. Fill in any missing combinations of ground truth & CpG-state calls
#    (i.e. where ground truth 'x' was never called as CpG-state 'y'),
#    add zero value:
    complete(ground_truth, CpG_State) %>%
    replace_na(list(call_rate = 0))
}
# Apply above function to generate mean CpG state calls on list of dataframes
mean_gt_list <- map(gt_list, fn_mean_calls)
```

# Accounting for the relative orientation of the ground truth with the reference sequence of synthetic spike-ins

For the synthetic spike-ins, there is no physical distinction at the molecular level between corresponding asymmetric states, e.g. 'MH' and 'HM'. Half the time, the same spike-in will go through the library prep in one orientation, and half the time, in the other:

  Orientation 1 ('HM')       Orientation 2 ('MH')
  Strand A  5'---HG---3'     Strand B  5'---MG---3'
  Strand B  3'---GM---5'     Strand A  3'---GH---5'
  
The distinction between HM and MH only arises when we create the reference sequence and must assign one strand of a duplex to be the 'plus' strand and the other to be the 'minus'. It is only then that we arbitrarily define whether the ground truth of this CpG site is an 'MH' or 'HM' state. 

Once we have made this arbitrary decision, though, the distinction between HM and MH matters. For example, if we choose HM to be the ground truth, rather than MH, then following correct alignment to the reference and base calling, it will be called as an HM state. If it were called as MH, this would reflect an error and should be counted as such. Furthermore, the HM state may have a different likelihood of being miscalled as HC than CH. This applies to any other asymmetric CpG state, but not to symmetric states: whether we define the ground truth as HM or MH, either state would have the same probability of being called as CC.

Thus, the HM and MH states can't simply be merged together and averaged, as we would not be able to distinguish correct HM calls from false MH calls, or separate the differing likelihood of HM being called as HC vs CH or other asymmetric states. 

In summary, there is no distinction at the molecular level between HM and MH. The distinction only arises with respect to alignment to the reference sequence. Thus, in the analysis below, both of these factors are accounted for; the ground truths are grouped by their two relative orientations w.r.t. the reference sequence (to reflect the arbitrary choice of the plus strand) and averaged accordingly (to reflect the fact that there is no distinction at the molecular level between the states).

```{r}
# Due to the definition of the plus strand of spike-in 'H2', the MH/MH state
# ground truth is labelled 'HM' and there is arbitrarily no ground truth for MH.
# To account for the points above, the MH ground truth must be added.
# (This was not necessary for any other asymmetric CpG state because both
# orientations happened to be covered when defining the reference sequences.)

# Add values for MH ground truth:
## Create vector with replicate names
## Create df with MH ground truth and CpG-state calls
MH_ground_truth <-
  data.frame(rep(c("MH"), times = 9),
             c("CC", "CM", "MC", "MM", "CH", "HC", "HH", "HM", "MH"),
             rep(c(NA), times = 9))
colnames(MH_ground_truth) <- c("ground_truth", "CpG_State", "call_rate")

## Define function to merge MH ground truth values with list
fn_merged_reps_with_MH <- function(df) {
  full_join(df,
            MH_ground_truth)  
}
# Apply above function to list
list_with_MH <- map(mean_gt_list, fn_merged_reps_with_MH)


# Define function to group ground truths and CpG-state calls
# by relative orientation to the reference:
fn_group_by_orient <- function(df) {
  df %>%
    mutate(
         GT_orient = case_when((ground_truth == "MC" |
                                  ground_truth == "HC" |
                                  ground_truth == "HM") ~ 1,
                               (ground_truth == "CM" |
                                  ground_truth == "CH" |
                                  ground_truth == "MH") ~ -1,
                               (ground_truth == "CC" |
                                  ground_truth == "MM" |
                                  ground_truth == "HH") ~ 0),
         CpG_State_call_orient = case_when((CpG_State == "MC" |
                                              CpG_State == "HC" |
                                              CpG_State == "HM") ~ 1,
                                           (CpG_State == "CM" |
                                              CpG_State == "CH" |
                                              CpG_State == "MH") ~ -1,
                                           (CpG_State == "CC" |
                                              CpG_State == "MM" |
                                              CpG_State == "HH") ~ 0),
         GT_form = case_when((ground_truth == "MC" |
                                ground_truth == "CM") ~ "MC/CM",
                             (ground_truth == "HC" |
                                ground_truth == "CH") ~ "HC/CH",
                             (ground_truth == "HM" |
                                ground_truth == "MH") ~ "HM/MH",
                             TRUE ~ ground_truth),
         CpG_State_call_form = case_when((CpG_State == "MC" |
                                            CpG_State == "CM") ~ "MC/CM",
                                         (CpG_State == "HC" |
                                            CpG_State == "CH") ~ "HC/CH",
                                         (CpG_State == "HM" |
                                            CpG_State == "MH") ~ "HM/MH",
                                         TRUE ~ CpG_State),
         group = abs(GT_orient + CpG_State_call_orient))
}

# Apply above function to list of dataframes
list_group_by_orient <- map(list_with_MH, fn_group_by_orient)


# Fn to determine call rates for grouped data, omitting MH to avoid NA values  
fn_group_call_rates <- function(df) {
  df %>%
    filter(ground_truth != "MH") %>%
    group_by(GT_form, CpG_State_call_form, group) %>%
    reframe(grouped_call_rate = mean(call_rate))
}
# Apply fn to list of dataframes
list_grouped_call_rates <- map(list_group_by_orient, fn_group_call_rates)
    
# Fn to merge grouped data with MH
fn_grouped_call_rates_all_ground_truth <- function(df1, df2) {
  left_join(df1, df2)
}

# Apply above fn
list_group_call_rates_all_gt <- Map(fn_grouped_call_rates_all_ground_truth, list_group_by_orient, list_grouped_call_rates)


# Fn to calculate mean ground truth -- CpG state combinations
fn_mean_grouped_call_rate <- function(df) {
  df %>%
    group_by(ground_truth, CpG_State) %>%
    reframe(grouped_call_rate = mean(grouped_call_rate))
}

# Apply fn to list
list_final_grouped_call_rates <- map(list_group_call_rates_all_gt, fn_mean_grouped_call_rate)


# Write call-rate data to file

## Rep 1 (JSH9):
final_grouped_call_rates_JSH9 <- list_final_grouped_call_rates[[1]] %>%
  write_tsv(paste0(call_rate_output, 
                   "final_grouped_call_rates_JSH9.tsv"))

## Rep 2 (JSH11)
final_grouped_call_rates_JSH11 <- list_final_grouped_call_rates[[2]] %>%
  write_tsv(paste0(call_rate_output, 
                   "final_grouped_call_rates_JSH11.tsv"))

## Reps merged 
final_grouped_call_rates_reps_merged <- list_final_grouped_call_rates[[3]] %>%
  write_tsv(paste0(call_rate_output, 
                   "final_grouped_call_rates_reps_merged.tsv"))
```

# Plots: call rate matrices
```{r}
## Specify order of CpG states
CpG_state_order <- c("CC", "MC", "CM", "MM", "HC", "CH", "HM", "MH", "HH")

# Fn to create call-rate matrix plot
fn_call_rate_matrix_plot <- function(df) {
  df %>%
    ggplot(aes(ground_truth, CpG_State, fill = grouped_call_rate)) +
    geom_tile() +
    theme_minimal(base_size = 10) +
    theme(legend.position = "none") +
    scale_fill_viridis(begin = 0.5, end = 1, alpha = 1) +
  #  scale_fill_gradient(low = "white", high = "blue") +
    scale_x_discrete(position = "top", limits = CpG_state_order) +
    scale_y_discrete(limits = rev(CpG_state_order)) +
    geom_text(aes(label = format(round(grouped_call_rate, digits = 2),
                               nsmall = 2), vjust = 0.5),
              size = 3) +
    xlab("True CpG state:") +
    ylab("CpG-state called (%)")
}

call_rate_matrix_rep1_JSH9 <- 
  fn_call_rate_matrix_plot(final_grouped_call_rates_JSH9) %>%
  ggsave(filename = paste0(call_rate_output, 
                "call_rate_matrix_rep1_JSH9.pdf"),
         width = 4, height = 3.5)

call_rate_matrix_rep2_JSH11 <- 
  fn_call_rate_matrix_plot(final_grouped_call_rates_JSH11) %>%
  ggsave(filename = paste0(call_rate_output, 
                "call_rate_matrix_rep2_JSH11.pdf"),
         width = 4, height = 3.5)

call_rate_matrix_reps_merged <- 
  fn_call_rate_matrix_plot(final_grouped_call_rates_reps_merged) %>%
  ggsave(filename = paste0(call_rate_output, 
                "call_rate_matrix_reps_merged.pdf"),
         width = 4, height = 3.5)
  
```

# Determining Pearson's correlation between call-rate matrices
```{r}
# Note that this was done without averaging corresponding asymmetric CpG states,
# as this would artificially inflate the correlation coefficient

pearsons_between_reps <-
  cor.test(list_group_call_rates_all_gt[[1]]$call_rate,
           list_group_call_rates_all_gt[[2]]$call_rate)

```