---
title: "enhancers"
author: "Jack Hardwick"
date: "2023-12-06"
output: html_document
---

# The following script performs the enhancer analysis, including:
* Average levels of CpG states at poised, primed, and active enhancers
* Profiles of CpG states at the above enhancer types
* Plots of the above

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
hmC_states <- c("HC+CH", "HM+MH", "HH")
non_hmC_states <- c("CC", "MM", "MC+CM")

# Load config file storing paths and experiment-specific variables
## Read config file where paths are defined
config <- read.table("config.txt", 
                     sep="=", 
                     comment.char="#", 
                     col.names=c("key", "value"))

## Convert contents to individual variables
config_list <- as.list(setNames(config$value, config$key)) %>%
  list2env(envir = .GlobalEnv)
```

### Obtaining coordinates for enhancers (primed, poised, and active)

The Excel file containing the enhancer coordinates: 
`./data/misc_files/1-s2.0-S1934590917300371-mmc2.xls` 
was downloaded from https://doi.org/10.1016/j.stem.2017.02.004.

# Loading enhancer coordinates
```{r}
fn_enhancers <- function(enh_type, sheet_no, output_filename){
data <- readxl::read_excel(paste0(misc_files, "1-s2.0-S1934590917300371-mmc2.xls"),
                     sheet = sheet_no,
                     col_types = "text")
sorted_data <- data[order(data$Chr, data$Start, data$End), ] %>%
  mutate(enh_type = enh_type) %>%
    write_tsv(paste0(misc_files, output_filename), col_names = FALSE)
}

# Split bivalent genes by Dppa2/4 status into separate bed files
active_enh <- fn_enhancers("active", 1, "active_enh.sorted.bed")
primed_enh <- fn_enhancers("primed", 2, "primed_enh.sorted.bed")
poised_enh <- fn_enhancers("poised", 3, "poised_enh.sorted.bed")

# Concatenate enhancer files, sort, and write to file
all_enh <- rbind(active_enh, primed_enh, poised_enh)
all_enh <- all_enh[order(all_enh$Chr, all_enh$Start, all_enh$End), ] %>%
  write_tsv(paste0(misc_files,
                   "all_enh.bed"), col_names = FALSE)
```

### Calculate mean lengths of active, primed, and poised enhancers
```{r}
# Load coords of all enhancers
all_enh <- read_tsv(paste0(misc_files, 
                           "all_enh.bed"), col_names = FALSE)

# Calculate length of each enhancer
all_enh_length <- all_enh %>%
  mutate(enh_length = X3 - X2) %>%
  group_by(X4) %>%
  reframe(mean_enh_length = mean(enh_length),
          median_enh_length = median(enh_length))

```
### Determine coordinates of CpG sites that overlap with all enhancers
```{zsh engine.opts='-i'}
conda activate doubleshot
source config.txt
mkdir -p $enh_output

# Combine enhancer coordinates into a single file and sort
cat ${misc_files}*_enh.sorted.bed | bedtools sort -i stdin > ${misc_files}all_enh.bed

# Find all CpG sites in both replicates that overlap with all enhancers
bedtools intersect -a ${misc_files}all_enh.bed -b ${intermed_files}quants_e14_5x_blacklist_filt_CpG_coords.bed -wa -wb > ${enh_output}enhancer_CpG_site_overlap_merged_reps.bed

```

### Shuffle enhancer files to compare enhancer data with random intervals
```{zsh engine.opts='-i'}
conda activate doubleshot

source config.txt

cd $enh_output

# Taking bins from replicate 1 (i.e. features in shuffled bed file must occur in regions covered by sequencing data). The choice of CC is arbitrary, as bed coordinates are the same for all binned CpG state files. Specifying seed integer enables shuffling to be exactly reproduced.
for file in ${misc_files}*_enh.sorted.bed; do
  bedtools shuffle -seed 927442958 -incl ${binned_data}bedgraph_merged/CC_quants_e14_filt_binned_all_chr.bedgraph -i $file -g ${misc_files}mm10.chrom.sizes.chr1_19XY.sorted > ${file%%.sorted.bed}_merged_reps_shuffle.bed
done

# Combine shuffled enhancer coordinates into a single files by replicate and sort
cat ${misc_files}*_merged_reps_shuffle.bed | bedtools sort -i stdin > ./all_enh_shuffle.merged_reps.bed

# Find all CpG sites in merged replicates that overlap with all enhancers
bedtools intersect -a ./all_enh_shuffle.merged_reps.bed -b ${intermed_files}quants_e14_5x_blacklist_filt_CpG_coords.bed -wa -wb > ./enhancer_shuffle_CpG_site_overlap.merged_reps.bed
```

### Determine average level of CpG states across each enhancer (also for grouped CpG states) and write to file
```{r}
# Import CpG site coordinates intersecting with all enhancer peaks
enh_CpG_overlap <-
  read_tsv(paste0(enh_output,
                  "enhancer_CpG_site_overlap_merged_reps.bed"),
           col_names = FALSE) %>% distinct()

# Import shuffled peaks
enh_shuffle_CpG_overlap <-
  read_tsv(paste0(enh_output,
                  "enhancer_shuffle_CpG_site_overlap.merged_reps.bed"),
           col_names = FALSE) %>% distinct()

# Name columns of shuffled peak files
colnames(enh_shuffle_CpG_overlap) <-
  c("feature_chr",
    "feature_start",
    "feature_end",
    "enh_type",
    "Chr",
    "start",
    "end")

# Label enhancer type with shuffle
enh_shuffle_CpG_overlap$enh_type <-
  paste0(enh_shuffle_CpG_overlap$enh_type, "_shuffled")

# Create column with unique feature ID for shuffled data
enh_shuffle_CpG_overlap <-
  enh_shuffle_CpG_overlap %>%
  mutate(feature_ID = paste0(enh_type,
                             sep = "_",
                             feature_chr, sep = "_", feature_start))

# Name columns for enhancer peaks
colnames(enh_CpG_overlap) <-
  c("feature_chr",
    "feature_start",
    "feature_end",
    "enh_type",
    "Chr",
    "start",
    "end")

# Create column with unique feature ID for enhancer peaks
enh_CpG_overlap <-
  enh_CpG_overlap %>%
  mutate(feature_ID = paste0(enh_type,
                             sep = "_",
                             feature_chr,
                             sep = "_",
                             feature_start))

# Import CpG-state info
CpG_states <-
  read_tsv(file = paste0(intermed_files, 
                         "quants_e14_5x_blacklist_filt.bed")) %>%
  tibble()

# Get CpG-state levels for CpG sites that overlap with enhancers
CpG_states_enh_overlap <- inner_join(CpG_states, enh_CpG_overlap)

# Get CpG-state levels for CpG sites that overlap with shuffled enhancers
CpG_states_enh_shuffle_overlap <-
  inner_join(CpG_states,
             enh_shuffle_CpG_overlap,
             by = join_by(Chr, start, end))

# Merge enhancer and shuffled data
CpG_states_enh_overlap <-
  rbind(CpG_states_enh_overlap, CpG_states_enh_shuffle_overlap)

# Remove unneeded data from memory
rm(CpG_states)

# Count number of CpG sites covered per enhancer in each replicate
CpG_states_enh_overlap_CpG_count <-
  CpG_states_enh_overlap %>%
  group_by(Chr, enh_type, feature_start, feature_end, feature_ID) %>%
  summarise(CpG_count = n_distinct(CpG_ID))

# Merge CpG count with df
CpG_states_enh_overlap <-
  left_join(CpG_states_enh_overlap,
            CpG_states_enh_overlap_CpG_count,
            by = join_by("Chr",
                         "feature_start",
                         "feature_end",
                         "enh_type",
                         "feature_ID"))

# Calculate mean pct of each CpG state across each enhancer
# and add zero to missing values (i.e. set mean level to 0
# where no CpG states were not detected within the enhancer).
CpG_states_enh_overlap_mean <- CpG_states_enh_overlap %>%
  group_by(Chr,
           enh_type,
           feature_start,
           feature_end,
           feature_ID,
           CpG_count,
           CpG_State) %>%
  reframe(mean_CpG_state_level = sum(pct) / CpG_count) %>%
  distinct() %>%
  complete(feature_ID, CpG_State, fill = list(mean_CpG_state_level = 0)) %>%
  distinct() %>%
  group_by(feature_ID) %>%
  fill(Chr,
       enh_type,
       feature_start,
       feature_end,
       CpG_count,
       .direction = "downup")

# Identify enhancers that do not contain a single covered CpG site
enhancer_covered <- CpG_states_enh_overlap_mean %>%
  group_by(feature_ID) %>%
  reframe(covered = as.logical(sum(mean_CpG_state_level))) %>%
  filter(covered == FALSE)

# Remove these non-covered enhancers from main df
CpG_states_enh_overlap_mean <-
  anti_join(CpG_states_enh_overlap_mean, enhancer_covered)

# Group corresponding asymmetric CpG states together
# (i.e. HC + CH, HM + MH, MC + CM)
# and calculate combined mean level across each enhancer:

## 1. Create new column 'CpG_state_group' for grouped asymmetric states
CpG_states_enh_overlap_mean <- CpG_states_enh_overlap_mean %>%
  mutate(CpG_state_group = case_when(CpG_State == "HC" |
                                       CpG_State == "CH" ~ "HC+CH",
                                     CpG_State == "HM" |
                                       CpG_State == "MH" ~ "HM+MH",
                                     CpG_State == "MC" |
                                       CpG_State == "CM" ~ "MC+CM",
                                     TRUE ~ CpG_State, ))

# Calculate sum of levels of grouped asymmetric states
CpG_states_enh_overlap_mean_grouped <-
  CpG_states_enh_overlap_mean %>%
  group_by(feature_ID,
           enh_type,
           CpG_count,
           CpG_state_group) %>%
  reframe(grouped_CpG_state_level = sum(mean_CpG_state_level))

# Write to file
CpG_states_enh_overlap_mean %>%
  write_tsv(paste0(enh_output,
                   "CpG_states_enh_overlap_by_CpG_state_reps_merged.tsv"))
CpG_states_enh_overlap_mean_grouped %>%
  write_tsv(paste0(enh_output,
                   "CpG_states_enh_overlap_by_CpG_group_reps_merged.tsv"))

# Compute summary stats for grouped CpG state
# levels for each enhancer type and write to file
CpG_states_enh_overlap_mean_grouped_summary_stats <-
  CpG_states_enh_overlap_mean_grouped %>%
  group_by(enh_type, CpG_state_group) %>%
  reframe(median = median(grouped_CpG_state_level),
          mean = mean(grouped_CpG_state_level))

# Compute summary stats for total hmC levels
# for each enhancer type and write to file
CpG_states_enh_overlap_total_hmC_summary_stats <-
  CpG_states_enh_overlap_mean_grouped %>%
  filter(CpG_state_group %in% hmC_states) %>%
  group_by(feature_ID) %>%
  mutate(total_hmC = sum(grouped_CpG_state_level)) %>%
  group_by(enh_type) %>%
  reframe(median_total_hmC = median(total_hmC),
          mean_total_hmC = mean(total_hmC))
```

### Plotting CpG state levels across enhancers
```{r}
# Import data
CpG_states_enh_overlap_mean <-
  read_tsv(paste0(enh_output,
                  "CpG_states_enh_overlap_by_CpG_state_reps_merged.tsv"))
CpG_states_enh_overlap_mean_grouped <-
  read_tsv(paste0(enh_output,
                  "CpG_states_enh_overlap_by_CpG_group_reps_merged.tsv"))

hmC_colours <- c("#3b528b", "#21908c", "#8fcfe0")

# Restructuring data for plot
CpG_states_enh_overlap_mean_grouped_restruct <-
  CpG_states_enh_overlap_mean_grouped %>%
  mutate(enh_type_shuffled = case_when(grepl("shuffle", enh_type) ~ "shuffled",
                                       TRUE ~ "true"))

CpG_states_enh_overlap_mean_grouped_restruct <-
  CpG_states_enh_overlap_mean_grouped_restruct %>%
  mutate(enh_type_full = case_when(grepl("active", enh_type) ~ "active",
                                   grepl("primed", enh_type) ~ "primed",
                                   grepl("poised", enh_type) ~ "poised"))

# Specify colours for plotting
enh_shuffle_colours <- RColorBrewer::brewer.pal(9, "Blues")[c(6, 4)]

# Plot grouped hmC states at enhancers (without shuffled comparison)
plot_grouped_hmC_enhancers <-
  CpG_states_enh_overlap_mean_grouped %>%
  filter(enh_type == "active" | enh_type == "poised" | enh_type == "primed") %>%
  filter(CpG_state_group %in% hmC_states) %>%
  ggplot(aes(x = factor(CpG_state_group,
                        levels = c("HC+CH", "HM+MH", "HH")),
             y = grouped_CpG_state_level,
             fill = factor(CpG_state_group,
                           levels = c("HC+CH", "HM+MH", "HH")))) +
  stat_boxplot(geom = "errorbar",
               position = position_dodge(0.9),
               width = 0.15) +
  geom_violin(size = 0.1,
              width = 0.5,
              linewidth = 0,
              position = position_dodge(0.9),
              scale = "width") +
  geom_boxplot(width = 0.5,
               alpha = 0.3,
               outlier.shape = NA,
               position = position_dodge(0.9),
               fill = "white") +
  coord_cartesian(ylim = c(0, 16)) +
  ggtitle("Levels of hmC states at enhancers (merged replicates)") +
  scale_fill_manual(values = hmC_colours) +
  scale_y_continuous(breaks = seq(0, 16, by = 2),
                     expand = c(0, 0)) +
  ylab("CpG-state level (%)") +
  xlab("CpG State") +
  facet_grid(~ factor(enh_type, levels = c("active", "primed", "poised"))) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        axis.line = element_line(linewidth = 0.4),
        axis.ticks = element_line(size = 0.4),
        axis.ticks.length = unit(0.1, "cm")) +
  guides(fill = guide_legend(title = ""))

ggsave(filename = paste0(enh_output, "plot_grouped_hmC_enhancers.pdf"), 
       plot = plot_grouped_hmC_enhancers,
       width = 6,
       height = 4.5)

# Plot grouped hmC states at enhancers and shuffled
plot_grouped_hmC_enh_shuffled <-
  CpG_states_enh_overlap_mean_grouped_restruct %>%
  filter(CpG_state_group %in% hmC_states) %>%
  ggplot(aes(x = factor(CpG_state_group,
                        levels = c("HC+CH", "HM+MH", "HH")),
             y = grouped_CpG_state_level,
             fill = factor(enh_type_shuffled,
                           levels = c("true", "shuffled")))) +
  stat_boxplot(geom = "errorbar",
               position = position_dodge(0.9),
               width = 0.15) +
  geom_violin(size = 0.1,
              linewidth = 0,
              position = position_dodge(0.9),
              scale = "width") +
  geom_boxplot(width = 0.5, alpha = 0.3,
               outlier.shape = NA,
               position = position_dodge(0.9)) +
  coord_cartesian(ylim = c(0, 15)) +
  ggtitle("Levels of hmC states at enhancers by replicate") +
  scale_fill_manual(values = enh_shuffle_colours) +
  scale_y_continuous(breaks = seq(0, 16, by = 2),
                     expand = c(0, 0)) +
  ylab("CpG-state level (%)") +
  xlab("CpG State") +
  facet_grid(~ enh_type_full) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        axis.line = element_line(linewidth = 0.4),
        axis.ticks = element_line(size = 0.4),
        axis.ticks.length = unit(.1, "cm")) +
  guides(fill = guide_legend(title = ""))
plot_grouped_hmC_enh_shuffled
ggsave(filename = paste0(enh_output, "plot_grouped_hmC_enh_shuffled.pdf"), 
       plot = plot_grouped_hmC_enh_shuffled,
       width = 6,
       height = 4.5)

non_hmC_state_levels_enh_plot <- 
  CpG_states_enh_overlap_mean_grouped_restruct %>%
  filter(CpG_state_group %in% non_hmC_states) %>%
  ggplot(aes(x = factor(CpG_state_group, levels = c("CC", "MM", "MC+CM")),
             y = grouped_CpG_state_level,
             fill = factor(enh_type_shuffled,
                           levels = c("true", "shuffled")))) +
  stat_boxplot(geom = "errorbar",
               position = position_dodge(0.9),
               width = 0.15) +
  geom_violin(size = 0.1,
              linewidth = 0,
              position = position_dodge(0.9),
              scale = "width") +
  geom_boxplot(width = 0.5,
               alpha = 0.3,
               outlier.shape = NA,
               position = position_dodge(0.9)) +
  ggtitle("Levels of CC, MM, and MC/CM at enhancers (merged replicates)") +
  ylab("CpG-state level (%)") +
  xlab("CpG State") +
  facet_grid(~ enh_type_full) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        axis.line = element_line(linewidth = 0.4),
        axis.ticks = element_line(size = 0.4),
        axis.ticks.length = unit(0.1, "cm")) +
  scale_fill_manual(values = enh_shuffle_colours) +
  scale_y_continuous(breaks = seq(0, 100, by = 10),
                     expand = c(0, 0)) +
  guides(fill = guide_legend(title = ""))

ggsave(filename = paste0(enh_output, "non_hmC_state_levels_enh_plot.pdf"), 
       plot = non_hmC_state_levels_enh_plot,
       width = 6,
       height = 4.5)


# Count number of enhancers in each group used for the plots 
# (i.e. those enhancers with CpG sites overlapping seq data)
fn_enh_count <- function(data, enh_group){
  data %>%
    filter(enh_type == enh_group) %>%
    dplyr::select(feature_ID) %>%
    unique() %>%
    nrow()
}

n_active_enh <- fn_enh_count(CpG_states_enh_overlap_mean_grouped, "active")
n_primed_enh <- fn_enh_count(CpG_states_enh_overlap_mean_grouped, "primed")
n_poised_enh <- fn_enh_count(CpG_states_enh_overlap_mean_grouped, "poised")

n_all_enh <- data.frame(n_active_enh, n_primed_enh, n_poised_enh) %>%
  write_tsv(file = paste0(enh_output,
                          "n_all_enh.tsv"))

```

# Total hmC at enhancers
```{r}
# Calculate total hmC per enhancer
CpG_states_enh_overlap_mean_grouped_total_hmC <-
  CpG_states_enh_overlap_mean_grouped_restruct %>%
  filter(CpG_state_group %in% hmC_states) %>%
  group_by(feature_ID, enh_type_shuffled, enh_type_full, CpG_count) %>%
  reframe(total_pct_hmC = sum(grouped_CpG_state_level))

# Plot total hmC across enhancers (true & shuffled)
plot_total_hmC_enhancers <-
  CpG_states_enh_overlap_mean_grouped_total_hmC %>%
  ggplot(aes(x = factor(enh_type_full,
                        levels = c("active", "primed", "poised")),
             y = total_pct_hmC,
             fill = factor(enh_type_shuffled,
                           levels = c("true", "shuffled")))) +
  stat_boxplot(geom = "errorbar",
               position = position_dodge(0.9),
               width = 0.15) +
  geom_violin(size = 0.1,
              width = 0.5,
              linewidth = 0,
              position = position_dodge(0.9),
              scale = "width") +
  geom_boxplot(width = 0.5,
               alpha = 0.3,
               outlier.shape = NA,
               position = position_dodge(0.9)) +
  coord_cartesian(ylim = c(0, 28)) +
  scale_y_continuous(breaks = seq(0, 35, by = 5),
                     expand = c(0, 0)) +
  ggtitle("Levels of total hmC states at enhancers (merged replicates)") +
#  scale_fill_manual(values = enh_colours) +
  #  facet_grid(~ enh_type_full) +
  ylab("CpG-state level (%)") +
  xlab("CpG State") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        axis.line = element_line(linewidth = 0.4),
        axis.ticks = element_line(size = 0.4),
        axis.ticks.length = unit(0.1, "cm")) +
  guides(fill = guide_legend(title = ""))

ggsave(filename = paste0(enh_output, "plot_total_hmC_enhancers.pdf"), 
       plot = plot_total_hmC_enhancers,
       width = 6,
       height = 4.5)

```

# Profiles of CpG states at enhancers
## First, prepare enhancer files, determining the centre of each enhancer
```{r}
# Load single bed file containing coordinates of all enhancer coordinates
all_enh <- read_tsv(paste0(misc_files,
                           "all_enh.bed"), col_names = FALSE)

# Determine midpoint of each enhancers, rounding up to nearest integer
all_enh_midpoints <- all_enh %>%
  mutate(start = ceiling(X2 + (X3 - X2) / 2), end = start + 1)

all_enh_midpoints %>%
  select(X1, start, end, X4) %>%
  write_tsv(paste0(misc_files, 
                   "all_enh_midpoints.bed"), col_names = FALSE)

# Split by enhancer type and write to file
all_enh_midpoints %>%
  filter(X4 == "active") %>%
  select(X1, start, end) %>%
  write_tsv(paste0(misc_files,
                   "active_enh_midpoints.bed"), col_names = FALSE)
```

## Split file containing enhancer midpoints by enhancer type then determine CpG-state profiles at each enhancer type using DeepTools
```{zsh engine.opts='-i'}
conda activate doubleshot

source config.txt

# Split by enhancer type
cd $misc_files
awk -F'\t' '{print > ("./" $4 "_enh_midpoints.bed")}' "all_enh_midpoints.bed" 
```

## Compute matrices for CpG states in DeepTools
```{zsh engine.opts='-i'}
conda activate doubleshot

source config.txt

# Create directories for deeptools matrices & plots
mkdir -p ${enh_output}deeptools_matrices_merged_reps/ \
    ${enh_output}deeptools_plots_merged_reps/

cd ${binned_data}bigwig_merged

# Primed enhancers: compute matrix for all states
for file in *.bw; do
  computeMatrix reference-point -p 8 \
  --referencePoint center \
  -S $file \
  -R ${misc_files}primed_enh_midpoints.bed \
  --upstream 8000 --downstream 8000 \
  --skipZeros \
  -o ${enh_output}deeptools_matrices_merged_reps/${file%%.bw}.primed_enh.mat.gz
done

# Active enhancers: compute matrix for all states
for file in *.bw; do
  computeMatrix reference-point -p 8 \
  --referencePoint center \
  -S $file \
  -R ${misc_files}active_enh_midpoints.bed \
  --upstream 8000 --downstream 8000 \
  --skipZeros \
  -o ${enh_output}deeptools_matrices_merged_reps/${file%%.bw}.active_enh.mat.gz
done

# Poised enhancers: compute matrix for all states
for file in *.bw; do
  computeMatrix reference-point -p 8 \
  --referencePoint center \
  -S $file \
  -R ${misc_files}poised_enh_midpoints.bed \
  --upstream 8000 --downstream 8000 \
  --skipZeros \
  -o ${enh_output}deeptools_matrices_merged_reps/${file%%.bw}.poised_enh.mat.gz
done
```

## Label matrices with file prefixes (needed for combining multiple states onto individual plots)
```{zsh engine.opts='-i'}
conda activate doubleshot

source config.txt

cd ${enh_output}deeptools_matrices_merged_reps

for file in *enh.mat.gz; do
  computeMatrixOperations relabel \
  -m $file \
  -o ${file%%.mat.gz}.relabel.mat.gz \
  --groupLabels ${file%%_binned*}
done

```

## Merge hmC matrices by replicate so that multiple CpG states can be combined on single plot
```{zsh engine.opts='-i'}
conda activate doubleshot

source config.txt

cd ${enh_output}deeptools_matrices_merged_reps

# Primed enhancers, all hmC
computeMatrixOperations rbind -m H*.grouped.primed_enh.relabel.mat.gz HH*.primed_enh.relabel.mat.gz -o hmC_primed_enh.merged_reps.mat.gz

# Primed enhancers, other states
computeMatrixOperations rbind -m MC*.grouped.primed_enh.relabel.mat.gz MM*.primed_enh.relabel.mat.gz CC*.primed_enh.relabel.mat.gz -o C_mC_primed_enh.merged_reps.mat.gz

# Active enhancers, all hmC
computeMatrixOperations rbind -m H*.grouped.active_enh.relabel.mat.gz HH*.active_enh.relabel.mat.gz -o hmC_active_enh.merged_reps.mat.gz

# Active enhancers, other states
computeMatrixOperations rbind -m MC*.grouped.active_enh.relabel.mat.gz MM*.active_enh.relabel.mat.gz CC*.active_enh.relabel.mat.gz -o C_mC_active_enh.merged_reps.mat.gz

# Poised enhancers, all hmC
computeMatrixOperations rbind -m H*.grouped.poised_enh.relabel.mat.gz HH*.poised_enh.relabel.mat.gz -o hmC_poised_enh.merged_reps.mat.gz

# Poised enhancers, other states, rep 1
computeMatrixOperations rbind -m MC*.grouped.poised_enh.relabel.mat.gz MM*.poised_enh.relabel.mat.gz CC*.poised_enh.relabel.mat.gz -o C_mC_poised_enh.merged_reps.mat.gz
```

## Plot peak profiles in DeepTools
```{zsh engine.opts='-i'}
conda activate doubleshot

source config.txt

# Primed enhancer plot
cd ${enh_output}deeptools_matrices_merged_reps
for file in hmC_*_enh.merged_reps.mat.gz; do
  plotProfile -m $file \
              -out ../deeptools_plots_merged_reps/${file%.mat.gz}.pdf \
              --numPlotsPerRow 1 \
              --refPointLabel enh.center \
              --yMax 9 \
              --yMin 0 \
              --plotWidth 6 \
              --plotHeight 10 \
              --plotTitle "Mod levels at enhancers (CpG-res)";
done

for file in C_mC_*_enh.merged_reps.mat.gz; do
  plotProfile -m $file \
              -out ../deeptools_plots_merged_reps/${file%.mat.gz}.pdf \
              --numPlotsPerRow 1 \
              --refPointLabel enh.center \
              --yMax 90 \
              --yMin 0 \
              --plotWidth 6 \
              --plotHeight 10 \
              --plotTitle "Mod levels at enhancers (CpG-res)";
done
```

